---
title: "CR Flex"
author: "Corey Scholes"
affiliation: "EBM Analytics"
version: 1.0
type: "website"
number-sections: true
number-depth: 3
date: "2026-Mar-10"
date-modified: "2026-Mar-10"

bibliography: AORAGov references.bib

editor:
  visual
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
    link-external-newwindow: true
    link-external-icon: true
    
    format:
  docx:
    toc: true
    number-sections: true

    
execute: 
  echo: false
  warning: false
  message: false
---

# Preamble

The following analysis is a report on the activity, quality and data contained in the AORA registry.

```{r}
#| label: load-packages

if (!require("pacman")) install.packages("pacman")
pacman::p_load(# Load required packages
  "patchwork",
  "nnet",
  "mgcv",
  "ggforce",
  "ggdag",
  "ggmap",
  "ggsurvfit",
  "geosphere",
  "dplyr",
  "flextable",
  "litedown",
  "grateful",
  "modelsummary",
  "quantreg",
  "readr",
  "knitr",
  "cardx",
  "forcats",
  "gargle",
  "googledrive",
  "googlesheets4",
  "openxlsx2",
  "tidyverse",
  "tidymodels",
  "lubridate",
  "consort",
  "gt",
  "gtsummary",
  "survival",
  "ggplot2",
  "ggdist",
  "ggfortify",
  "mice",
  "marginaleffects",
  "naniar",
  "quantreg",
  "broom",
  "epoxy",
  "glue",
  "broom.helpers",
  "stringr"
  )
  
  
```

Access to the AORA datasets was pre-authorised.

```{r}
#| label: auth2
#| echo: false

options(
  gargle_oauth_cache = ".secrets",
  gargle_oauth_email = TRUE
)

googledrive::drive_auth(
  cache = ".secrets", 
  email = TRUE
  )
```

A function was generated to retrieve files to call on later in the analysis for processing data imports.

```{r}
#| label: folder
#| echo: false

base_folder_id1 <- "10ThtWn4-IIOuTYPbNrhJkL6yBuPQjk5S"



```

```{r}
get_specific_snapshot <- function(folder_name, base_folder_id = base_folder_id1) {
  tryCatch({
    # Check if the folder exists in the base directory
    folder <- googledrive::drive_ls(as_id(base_folder_id), pattern = paste0("^", folder_name, "$"))
    
    if(nrow(folder) == 0) {
      stop(paste("Folder", folder_name, "not found"))
    }
    
    # Find the snapshot file in the specified folder
    snapshot_file <- googledrive::drive_ls(
      folder$id, 
      pattern = "Registry data snapshot\\.xlsx$"
    )
    
    if(nrow(snapshot_file) == 0) {
      stop("No snapshot file found in specified folder")
    }
    
    # Return both pieces of information as a list
    return(list(
      snapshot = snapshot_file,
      folder_name = folder$name
    ))
    
  }, error = function(e) {
    stop(paste("Error finding specified snapshot:", e$message))
  })
}
```

Data was retrieved from live database tables. Source files were specified and stored as global variables to call on in further functions.

```{r}
#| label: live-table
#| echo: false

SheetIDs <- list(
DbSS = "https://docs.google.com/spreadsheets/d/1fhW1LJbsdvkA7u-z1IGgI2qHwiOWXP6RMRH8-cgPXU8/edit",
IntHistory = "https://docs.google.com/spreadsheets/d/1YBYcxJtRIoH0rD87bWCBvIFiKQp-b-JmklWPOh_-wDY/edit"
)

# Treatments Masterlist - Timestamp is > Registry Start date (gsuite)

```

```{r}
#| label: read-tables

# Authenticate for sheets using the same token
gs4_auth(token = drive_token())



ComplicTable <- googlesheets4::read_sheet(
  ss = SheetIDs$DbSS,
  sheet = "Complications",
  col_names = TRUE,
  col_types = "TccDccDc"
  )


#To match to acctData
PatientTable <- googlesheets4::range_read(
  ss = SheetIDs$DbSS,
  sheet = "Patient",
  range = "A7:P",
  col_names = FALSE,
  col_types = "DccccDcccDciccci"
  )


Patient_Col <- c(
  "PatientCreationDate",
  "PatientID",
  "LastName",
  "FirstName",
  "AlternateID",
  "DateOfBirth",
  "Sex",
  "RegistryStatus",
  "RegistryStatusNotes",
  "DateRegistryStatus",
  "NotificationMethod",
  "NoTreatmentRecords",
  "Email",
  "Phone",
  "PostCode",
  "TrueNoTreatments"
)

colnames(PatientTable) <- Patient_Col


MonitorPlan <- googlesheets4::range_read(
  ss = SheetIDs$DbSS,
  sheet = "Monitor Module",
  range = "A1:O",
  col_names = TRUE,
  col_types = "ccccDTtincTccii"
  )


MonitorLog <- googlesheets4::range_read(
  ss = SheetIDs$DbSS,
  sheet = "MonitorModuleLog",
  range = "A1:K",
  col_names = TRUE,
  col_types = "Tcicccccccc"
  )

MonitorResponse <- googlesheets4::range_read(
  ss = SheetIDs$DbSS,
  sheet = "Monitoring",
  range = "A1:Q",
  col_names = TRUE,
  col_types = "TccccccccccnccDic"
  )

IntHistory <- googlesheets4::range_read(
  ss = SheetIDs$IntHistory,
  sheet = "AORA - Medical History_RawData", 
  range = "A4:CO",
  col_names = FALSE
  )

IntHistoryCol <- googlesheets4::range_read(
  ss = SheetIDs$IntHistory,
  sheet = "AORA - Medical History_RawData", 
  range = "A3:CO3",
  col_names = FALSE
  )

 colnames(IntHistory) <- IntHistoryCol

```

## Retrieve MMM

```{r}

# ── 1. Download the POA 2021 × MMM 2023 concordance ──────────────────────────
# Direct xlsx from data.gov.au (Dept of Health and Aged Care, CC BY 2.5 AU)
# https://www.data.gov.au/data/dataset/postal-areas-2021-by-modified-monash-model-2023
 
mmm_url <- "https://data.gov.au/data/dataset/9c4e8fab-e9c3-4728-80a8-e59643f258a5/resource/04ed31ba-d48b-437c-9279-f4b788b9dd90/download/poa_21-mmm_23.xlsx"

tmp <- tempfile(fileext = ".xlsx")
download.file(mmm_url, destfile = tmp, mode = "wb")

 concordance_raw <- openxlsx2::wb_to_df(tmp, start_row = 8, start_col = 2, skip_empty_cols = TRUE)

# ── 2. Inspect and clean the concordance ─────────────────────────────────────
# Preview structure first — column names may vary; check with:
 glimpse(concordance_raw)

MMMPostcode <- concordance_raw |>
  rename_with(str_to_lower) |>
  rename(
    postcode  = "poa 2021 code",
    mmm_code  = "mmm 2023"
  ) |>
  summarise(
    mmm_code       = max(mmm_code, na.rm = TRUE),
    mmm_ambiguous  = n() > 1,
    .by = postcode
  )  |>
  mutate(
    postcode = as.character(postcode),
    mmm_code = as.integer(mmm_code),
    mmm_label = case_when(
      mmm_code == 1 ~ "Metropolitan",
      mmm_code == 2 ~ "Regional centre",
      mmm_code == 3 ~ "Large rural town",
      mmm_code == 4 ~ "Medium rural town",
      mmm_code == 5 ~ "Small rural town",
      mmm_code == 6 ~ "Remote community",
      mmm_code == 7 ~ "Very remote community",
      .default      = NA_character_
    )
  ) |>
  select(postcode, mmm_code, mmm_label, mmm_ambiguous)

```

```{r}


IntHistory1 <- IntHistory |> dplyr::select(
  FormIDMedicalHistory,
  EducationLevel,
  SmokingStatus,
  AlcoholConsumption,
  ActivityDuration,
  SymptomOnset,
  SymptomDuration
) |> rename(
  FormID = "FormIDMedicalHistory"
) |> distinct(
  FormID,
  .keep_all = TRUE
  ) |> mutate(
    TreatmentID = str_extract(FormID, "^\\d+\\.\\d+"),
    SymptomDuration1 = if_else(
      as.numeric(str_extract(SymptomDuration, "\\d+")) >1000,
      NA,
      as.numeric(str_extract(SymptomDuration, "\\d+"))
  )) 

```

```{r}

KneeIntraop <- googlesheets4::range_read(
  ss = SheetIDs$DbSS,
  sheet = "Knee Arthritis Treatments", 
  range = "A2:X",
  col_names = TRUE, 
  col_types = "Tccccccccccccccccccccncc"
  )
```

A static registry snapshot was retrieved and formatted based on the fixed date of preparation of the snapshot (01-Dec-2025).

```{r}

# Authenticate for sheets using the same token
gs4_auth(token = drive_token())

# To get a snapshot from a specific folder (e.g., "20230415")
specific_snapshot <- get_specific_snapshot("20251130")


```

```{r}

temp_file1 <- tempfile(fileext = ".xlsx")
drive_download(
  file = specific_snapshot$snapshot$id,
  path = temp_file1,
  overwrite = TRUE
)

# Correction to reset back to excel origin
DaysDiff <- as.numeric(as.duration(interval(ymd("1899-12-30"), ymd("1970-01-01"))),"days")


Snapshot <- openxlsx2::wb_to_df(
  temp_file1,
  sheet = "Combined",
  colNames = TRUE,
  detectDates = TRUE
  ) |> mutate(
    PatientID = stringr::str_split_i(TreatmentID,"\\.",1),
     SurgicalTreatment2 = case_when(
       stringr::str_detect(str_to_lower(EBMAComment),"surgery recommended") &
       stringr::str_detect(str_to_lower(SurgicalTreatment), "non") ~ "Surgery recommended",
       stringr::str_detect(str_to_lower(SurgicalTreatment), "non", negate = TRUE) ~ "Surgical",
       stringr::str_detect(str_to_lower(SurgicalTreatment), "non") ~ "Non-Surgical"
  )
  ) |> unite(
  col = "CombID",
  sep = ".",
  c("PatientID","AffectedSide"),
  remove = FALSE
)




STROBEInput <- openxlsx2::wb_to_df(
  temp_file1,
  sheet = "Strobe_Input",
  colNames = TRUE,
  detectDates = TRUE
  )

```

# Context

The Arthroplasty Outcomes in Regional Australia (AORA) registry has been in operation since June 2020, collecting KneeArthritis and HipArthritis presentations to a private clinic in Grafton, NSW. The setup of the registry is detailed on the [registry wiki](https://sites.google.com/ebma.com.au/aora-registry-wiki/home). The registry started with two surgeons contributing with one surgeon stopping participation in December 2022.

# Recruitment Flow

Flowcharts as per STROBE [@Vandenbroucke2007] and RECORD [@Benchimol2015] guidelines were generated for treatments enrolled into the Registry. Followup was set to eligibility at 6months followup.

Need to identify GMKSubset to flag in STROBEInput for flowchart

```{r}

CurrentDate <- as.character("20251130")


STROBEFlow <- STROBEInput |> left_join(
  KneeIntraop |> dplyr::select(
    -Timestamp
  ),
  join_by(TreatmentID == "Treatment ID")
) |> dplyr::left_join(# Join with relevant columns from Snapshot
    Snapshot |> 
      dplyr::select(
        TreatmentID,
        CombID,
        EligibleAtPreop,
        EligibleAtx6months,
        EligibleAtx6months,
        SurgicalTreatment2
      ) |> 
      # Convert all date columns to date format
      dplyr::mutate(
        across(contains("Date"), lubridate::ymd)
      ),
    by = "TreatmentID"
  ) |> dplyr::mutate(
    TreatmentStatus2 = if_else(
          TreatmentStatus == "No further followup" & is.na(DateTreatment),
          "Archived",
          TreatmentStatus
  )
  ) |> dplyr::mutate(
    exclusion1 = case_when(
    RegistryCohortName != "KneeArthritis" ~ "Diagnosis not of interest",
    (SurgicalTreatment2 != "Surgical" | is.na(SurgicalTreatment2) | str_detect(TreatmentStatus,"Pending")) & RegistryCohortName == "KneeArthritis" ~ "Not a surgical treatment",
     SurgicalTreatment2 == "Surgical" & RegistryCohortName == "KneeArthritis" & stringr::str_detect(RegistryStatus,"Opt-out", negate = TRUE) & TreatmentType != "Primary" ~ "Non-primary procedure",
    (SurgicalTreatment2 != "Surgical" | !is.na(SurgicalTreatment2)) & RegistryCohortName == "KneeArthritis" & stringr::str_detect(RegistryStatus,"Opt-out") ~ "Patient Opt-Out",
    #TreatmentStatus2 == "Archived" ~ "Record archived",
    .default = NA_character_
  ),
  induction = if_else(
    is.na(exclusion1),
    TreatmentID,
    NA_character_
  ),
Hardware1 = case_when(
  is.na(induction) ~ NA_character_,
  !is.na(induction) & is.na(`Femur Brand`) ~ "No data",
  !is.na(induction) & !is.na(`Femur Brand`) & str_detect(str_to_lower(`Femur Brand`),"gmk|gmc|sph*|medacta") ~ "Brand",
  !is.na(induction) & !is.na(`Femur Brand`) & str_detect(str_to_lower(`Femur Brand`),"gmk|gmc|sph*|medacta", negate = TRUE) ~ "Incorrect hardware"
  ),
Hardware2 = case_when(
  is.na(induction) ~ NA_character_,
  !is.na(induction) & is.na(`Bearing Type`) ~ "No data",
  !is.na(induction) & !is.na(`Bearing Type`) & str_detect(str_to_lower(`Bearing Type`),"medial") ~ "Type",
  !is.na(induction) & !is.na(`Femur Brand`) & str_detect(str_to_lower(`Femur Brand`),"medial", negate = TRUE) ~ "Incorrect Type"
  ),
Hardware = str_c(Hardware1,Hardware2, sep = ""),
exclusion2 = case_when(
  !is.na(induction) & stringr::str_detect(Hardware, "Incorrect") ~ "Not hardware of interest",
  !is.na(induction) & Hardware == "No dataNo data" ~ "No data entered"
),
subjid_dosed = if_else(
    is.na(exclusion1) & is.na(exclusion2),
    TreatmentID,
    NA_character_
),
arm = case_when(
is.na(exclusion1) & is.na(exclusion2) & DateTreatment <= ymd("2022-Nov-02") ~ "Pre",
is.na(exclusion1) & is.na(exclusion2) & DateTreatment >= ymd("2022-Nov-02") ~ "Post"
),
arm = forcats::fct(
  arm,
  levels = c("Pre","Post")
),
followup = if_else(
  !is.na(arm),
  TreatmentID,
  NA_character_
),
lost_followup = if_else(
  EligibleAtx6months != "Yes",
  "Not eligible for followup",
  NA_character_
),
mitt = if_else(
    !is.na(followup) & is.na(lost_followup),
    TreatmentID,
    NA_character_
)
) |> dplyr::rename(
  trialno = "TreatmentID",
  arm3 = "RegistryCohortName"
)

```

```{r}

STROBEReview <- STROBEFlow |> group_by(arm) |> summarise(CohortCount = n()) |> arrange(desc(CohortCount))

```

```{r}
#| label: fig-strobe-cohort1
#| fig-cap: "Flowchart of extraction and followup of surgical sample from the Registry"


STROBEPlot1 <- consort_plot(
  data = STROBEFlow,
 orders = c(
   trialno = "Population",
   exclusion1 = "Ineligible for TKA",
   induction = "Surgical pool",
   exclusion2 = "Excluded",
   subjid_dosed = "Total cohort",
   arm = "Grouping",
   lost_followup = "Not eligible for followup",
   mitt = "Followup Analysis"
   ),
 side_box = c(
   "exclusion1",
   "exclusion2",
   "lost_followup"
   ),
 allocation = "arm",
 cex = 0.6
)

knitr::knit_print(STROBEPlot1)

```

Cumulative recruitment over time was plotted from Registry inception to the present.

```{r}
#| label: fig-progress
#| fig-cap: "Cumulative cohort recruitment within the registry from inception to present"

target_date <- lubridate::ymd("2022-Nov-02")  # Set your date here

RecordCount <- dplyr::summarise(
  STROBEFlow |>  dplyr::filter(
    !is.na(followup)
    ),
  Count = n(),
  .by = PatientCreationDate
) |> arrange(
  PatientCreationDate
) |> dplyr::mutate(
  CumulativeCount = cumsum(Count),
  PatientCreationNum = as.numeric(PatientCreationDate)
  ) |> filter(
    !(PatientCreationDate < ymd("20160701"))
   
  )

cumulative_plot <- ggplot(RecordCount, aes(x = PatientCreationDate, y = CumulativeCount)) +
  geom_line(color = "#2c3e50", linewidth = 1) +
  # Add the spline fit with confidence interval
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), 
              color = "#e74c3c", linewidth = 1.2, se = TRUE, fill = "#e74c3c", alpha = 0.2) +
  geom_point(color = "#3498db", size = 1) +
  
  geom_vline(xintercept =  target_date,
             linetype = "dashed", color = "#27ae60", linewidth = 0.8) +
  annotate("text", x = target_date, y = Inf,
           label = "Introduction of Flex bearing", vjust = 1.5, hjust = -0.1,
           color = "#27ae60", fontface = "bold", size = 3.5) +
  # Add labels and title
  labs(
    x = "Date",
    y = "Cumulative Number of Cases",
    #caption = "Data source: STROBEFlow3"
  ) +
  # Improve the theme and formatting
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

# Print the plot
knitr::knit_print(cumulative_plot)


```



# Baseline and Demographics

```{r}
GMKMaster <- Snapshot |> filter(
  TreatmentID %in% STROBEFlow$followup  
) |> left_join(
  IntHistory1 |> dplyr::select(
    TreatmentID,
    EducationLevel:SymptomOnset,
    SymptomDuration1
  ),
  by = "TreatmentID"
) |> left_join(
  MMMPostcode,
  join_by(Postcode == postcode),
  relationship = "many-to-one"
) |> mutate(
  across(
      ComorbHeartDisease_Preop:ComorbOtherMedicalCondition_Preop,
      ~ case_when(
        str_detect(.x, "not") ~ 0,
        str_detect(.x, "not|activities|treatment", negate = TRUE) ~ 1,
        str_detect(.x, "treatment") ~ 2,
        str_detect(.x, "activities")~ 3
      )
    ),
    SACQ_score = if_else(
      rowSums(!is.na(pick(ComorbHeartDisease_Preop:ComorbOtherMedicalCondition_Preop))) == 0L,
      NA_integer_,
      rowSums(
        pick(ComorbHeartDisease_Preop:ComorbOtherMedicalCondition_Preop),
        na.rm = TRUE
    )
    ),
  across(
      MODEMQ1_Preop:MODEMQ6_Preop,
      ~ case_when(
        str_detect(.x, "Extremely") ~ 4,
        str_detect(.x, "Very") ~ 3,
        str_detect(.x, "Somewhat") ~ 2,
        str_detect(.x, "Slightly") ~ 1,
        str_detect(.x, "Not.likely") ~ 0,
        str_detect(.x, "applicable") ~ NA,
        is.na(.x) ~ NA
      )
    ),
    MODEMExp = if_else(
      rowSums(!is.na(pick(MODEMQ1_Preop:MODEMQ6_Preop))) == 0,
      NA_integer_,
      rowSums(
        pick(MODEMQ1_Preop:MODEMQ6_Preop),
        na.rm = TRUE
    )
    ),
  
CRTiming = case_when(
DateTreatment <= ymd("2022-Nov-02") ~ "Pre",
DateTreatment >= ymd("2022-Nov-02") ~ "Post"
),
CRTiming = forcats::fct(
  CRTiming,
  levels = c("Pre","Post")
),
)


```

```{r}

BearingPlot <- GMKMaster |>
  ggplot(aes(x = DateTreatment, y = CRTiming)) +
  geom_point(alpha = 0.6, size = 2) +
  labs(
    title = "Cruciate retaining use over time",
    x = "Date",
    y = "Before or after first CR use"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

knitr::knit_print(BearingPlot)
```

```{r}
#| label: tbl-crflex-demographics
#| tbl-cap: "Summary of demographics comparing pre vs post cr introduction"

Tableflexdemo <- gtsummary::tbl_summary(
  GMKMaster |> dplyr::select(
    CRTiming,
    AgeAtTreatment,
    Sex,
    EducationLevel,
    mmm_label,
    SACQ_score,
    MODEMExp,
    KOOS12_Pain_TotalScore_Preop,
    VR12_Physical_TotalScore_Preop,
    VR12_Mental_TotalScore_Preop,
    TreatmentStatus,
    DateTreatment
  ),
  by = "CRTiming",
  missing = "no",
  statistic = list(
    DateTreatment ~ "{min} - {max}"
  )
) |>
  add_overall()

knitr::knit_print(Tableflexdemo)
```

# Complications

```{r}
GMKComp <- GMKMaster |> dplyr::select(
  TreatmentID,
  CRTiming,
  DateTreatment,
  TreatmentStatus,
  TreatmentStatusNotes,
  DateStatusChange,
  ComplicationOccurrence,
  DateOfOccurrence,
  ComplicationNature,
  ReoperationProcedure,
  ReoperationProcedureDate
) |> mutate(
  ComplicationOccur2 = if_else(
    is.na(ComplicationOccurrence),
    "No",
    ComplicationOccurrence
  ),
  OccurDelay =  as.numeric(as.duration(interval(ymd(DateTreatment), ymd(DateOfOccurrence))),"days"),
  Followup = if_else(
    is.na(DateStatusChange),
    as.numeric(as.duration(interval(ymd(DateTreatment), ymd(CurrentDate))),"days"),
    as.numeric(as.duration(interval(ymd(DateTreatment), ymd(DateStatusChange))),"days")
  )
)


```

```{r}
#| label: tbl-comp
#| tbl-cap: "Summary of complications incidence after surgery separated by group"

TableGMKComp <- gtsummary::tbl_summary(
  GMKComp |> dplyr::select(
    CRTiming,
    ComplicationOccur2,
    OccurDelay,
    Followup
  ),
  by = "CRTiming",
  missing = "no"
) |>
  modify_footnote_body(
    footnote = "Days",
    columns = "label",
    rows = variable == c("OccurDelay","Followup") & row_type == "label"
  ) |> add_overall()

knitr::knit_print(TableGMKComp)


```

Adverse events were detected in 16% of surgical cases performed by SM over the life of the registry as shown in @tbl-comp, with an average detection with 30 days of surgery.

Write complications table for SM review

```{r}
#| label: write-complic
#| execute: false

# Authenticate for sheets using the same token
gs4_auth(token = drive_token())


# CRFlexComp <- googlesheets4::gs4_create(
#   "AORA_CRFlexComp",
#   sheets = "Complications"
# )

sheet_write(
  GMKComp |> dplyr::select(
    -ComplicationOccurrence
  ) |> filter(
    ComplicationOccur2 != "No"
  ),
  ss = "https://docs.google.com/spreadsheets/d/13AcKTFGWuz_qZw3mNBRKSrta0CIEflUDZPohfLF26Cg?usp=drive_fs",
  sheet = "Complications"
)

```




# Patient reported outcomes

```{r}

CRKOOSADL <- GMKMaster |> ggplot(
  aes(y = KOOS12_DailyLiving_TotalScore_6months, x = CRTiming)) +
  geom_dots() +
  stat_halfeye(
    alpha = 0.5,  # Transparency for overlap visibility
    position = "identity",  # Overlay the distributions
    na.rm = TRUE,
    scale = 0.9  # Slightly scale down to avoid too much overlap
  )

knitr::knit_print(CRKOOSADL)
```


Ran a multinomial logistic model to assess differences in kneeling difficulty at 6months followup before vs after introduction of flex bearing

```{r}

ancova_tbl <- function(data, outcome, baseline_cov, label) {
  
  formula <- as.formula(paste(outcome, "~ CRTiming+", baseline_cov, "+ AgeAtTreatment + Sex"))
  
  lm(formula, data = data) |>
    tbl_regression(
      include   = "CRTiming",
      intercept = FALSE,
      label     = list(CRTiming ~ label)
    ) |>
    bold_p(t = 0.05) |>
    bold_labels()
}
```

```{r}

tbl_dl <- ancova_tbl(
  data         = GMKMaster,
  outcome      = "KOOS12_DailyLiving_TotalScore_6months",
  baseline_cov = "KOOS12_DailyLiving_TotalScore_Preop",
  label        = "Daily Living"
)

tbl_pain <- ancova_tbl(
  data         = GMKMaster,
  outcome      = "KOOS12_Pain_TotalScore_6months",
  baseline_cov = "KOOS12_Pain_TotalScore_Preop",
  label        = "Pain"
)

tbl_total <- ancova_tbl(
  data         = GMKMaster,
  outcome      = "KOOS12_Summary_TotalScore_6months",
  baseline_cov = "KOOS12_Summary_TotalScore_Preop",
  label        = "Total"
)

# ancova_KOOS12ADL <- lm( ~ CRTiming + KOOS12_DailyLiving_TotalScore_Preop + AgeAtTreatment + Sex, data = )
```

```{r}
#| label: tbl-proms
#| tbl-cap: "ANCOVA: 6month KOOS scores adjusted for baseline"


tbl_stacked <- tbl_stack(
  tbls         = list(tbl_dl, tbl_pain, tbl_total),
  group_header = c("Daily Living", "Pain", "Total")
) |>
  modify_table_body(
    ~ .x |>
        filter(row_type == "level" & label == "Post") |>
        mutate(
          label         = groupname_col,
          groupname_col = "",
          row_type      = "tbl_summary-str:header-nlevel"
        )
  ) |> modify_footnote(
    update = everything() ~ "β = adjusted mean difference (Treatment vs Control); adjusted for baseline subscale score."
  )



knitr::knit_print(tbl_stacked)

```


```{r}
#| label: tbl-proms2
#| tbl-cap: "Summary of 6month KOOS scores adjusted for baseline, age and sex"

# ── 1. Fit models ────────────────────────────────────────────────────────────

fit_ancova <- function(data, outcome, baseline_cov) {
  as.formula(
    paste(outcome, "~ CRTiming +", baseline_cov, "+ AgeAtTreatment + Sex")
  ) |>
    lm(data = data)
}

fit_dl    <- fit_ancova(GMKMaster, "KOOS12_DailyLiving_TotalScore_6months", "KOOS12_DailyLiving_TotalScore_Preop")
fit_pain  <- fit_ancova(GMKMaster, "KOOS12_Pain_TotalScore_6months",        "KOOS12_Pain_TotalScore_Preop")
fit_total <- fit_ancova(GMKMaster, "KOOS12_Summary_TotalScore_6months",     "KOOS12_Summary_TotalScore_Preop")


# ── 2. Extract avg marginal predictions per model ───────────────────────────

extract_preds <- function(fit, col_label) {
  
  # Subset to only the columns used in the model
  model_vars <- all.vars(formula(fit))
  newdata    <- fit$model |> select(all_of(model_vars))
  
  invisible(avg_predictions(fit, variables = "CRTiming", newdata = newdata) |>
    as_tibble() |>
    mutate(
      cell = glue("{round(estimate, 1)} ({round(conf.low, 1)}, {round(conf.high, 1)})")
    ) |>
    select(CRTiming, cell) |>
    rename(!!col_label := cell)
  )
}

tbl_dl_pred    <- extract_preds(fit_dl,    "Daily Living")
tbl_pain_pred  <- extract_preds(fit_pain,  "Pain")
tbl_total_pred <- extract_preds(fit_total, "Total")


# ── 3. Merge and render gt table ─────────────────────────────────────────────

tbl_combined <- tbl_dl_pred |>
  left_join(tbl_pain_pred,  by = "CRTiming") |>
  left_join(tbl_total_pred, by = "CRTiming") |>
  rename(Group = CRTiming)

tbl_combined |>
  gt() |>
  tab_header(
    title    = "Model-Predicted KOOS-12 Scores at 6 Months",
    subtitle = "Adjusted marginal means (95% CI) by CRTiming group"
  ) |>
  tab_spanner(
    label   = "KOOS-12 Subscale",
    columns = c("Daily Living", "Pain", "Total")
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = Group)
  ) |>
  tab_footnote(
    footnote  = "Adjusted marginal means from ANCOVA, computed at observed covariate values (age, sex, baseline score).",
    locations = cells_title(groups = "subtitle")
  ) |>
  opt_row_striping() |>
  opt_table_font(font = google_font("Source Sans Pro"))

```



```{r}
#| label: fig-kneeling
#| fig-cap: "Response patterns of kneeling and daily living difficulty at 6months before vs after introduction of cr bearing"
#| fig-width: 8
#| fig-height: 7
#| dpi: 300
#| fig-align: center

# ── Shared diverging palette (None = blue, Extreme = red) ────────────────────
# RdYlBu reversed so severity reads red → yellow → blue
pal_5 <- rev(RColorBrewer::brewer.pal(5, "RdYlBu"))
names(pal_5) <- c("Extreme", "Severe", "Moderate", "Mild", "None")

pal_4 <- rev(RColorBrewer::brewer.pal(4, "RdYlBu"))
names(pal_4) <- c("Severe", "Moderate", "Mild", "None")

# ── Kneeling ──────────────────────────────────────────────────────────────────
KneelPlotData <- GMKMaster |>
  filter(!is.na(Kneeling_6months)) |>
  mutate(
    Kneeling_6months = forcats::fct(
      Kneeling_6months,
      levels = c("Extreme", "Severe", "Moderate", "Mild", "None")
    )
  ) |>
  count(CRTiming, Kneeling_6months) |>
  group_by(CRTiming) |>
  mutate(
    proportion = n / sum(n),
    percentage = round(proportion * 100, 1)
  ) |>
  ungroup()

KneelPlot <- ggplot(KneelPlotData, aes(x = CRTiming, y = proportion, fill = Kneeling_6months)) +
  geom_col(position = "fill") +
  geom_text(
    aes(label = paste0(percentage, "%")),
    position = position_fill(vjust = 0.5),
    colour = "white",
    size = 2.8
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = pal_5) +
  labs(x = "CRTiming", y = "Proportion", fill = "Kneeling difficulty") +
  theme_minimal()

# ── ADL data prep ─────────────────────────────────────────────────────────────
ADLPlotData <- GMKMaster |>
  dplyr::select(
    TreatmentID, CRTiming,
    contains("DailyLiving") & contains("6months")
  ) |>
  mutate(
    across(
      c(KOOS12DailyLivingQ1_6months, KOOS12DailyLivingQ2_6months, KOOS12DailyLivingQ3_6months, KOOS12DailyLivingQ4_6months),
      ~ forcats::fct(., levels = c("Extreme", "Severe", "Moderate", "Mild", "None"))
    )
  )

make_adl_data <- function(data, var) {
  data |>
    count(CRTiming, {{ var }}) |>
    filter(!is.na({{ var }})) |>
    group_by(CRTiming) |>
    mutate(
      proportion = n / sum(n),
      percentage = round(proportion * 100, 1)
    ) |>
    ungroup()
}

ADLPlotData1 <- make_adl_data(ADLPlotData, KOOS12DailyLivingQ1_6months)
ADLPlotData2 <- make_adl_data(ADLPlotData, KOOS12DailyLivingQ2_6months)
ADLPlotData3 <- make_adl_data(ADLPlotData, KOOS12DailyLivingQ3_6months)
ADLPlotData4 <- make_adl_data(ADLPlotData, KOOS12DailyLivingQ4_6months)

make_adl_plot <- function(data, fill_var, legend_title, palette) {
  ggplot(data, aes(x = CRTiming, y = proportion, fill = {{ fill_var }})) +
    geom_col(position = "fill") +
    geom_text(
      aes(label = paste0(percentage, "%")),
      position = position_fill(vjust = 0.5),
      colour = "white",
      size = 2.8
    ) +
    scale_y_continuous(labels = scales::percent_format()) +
    scale_fill_manual(values = palette) +
    labs(x = "CRTiming", y = "Proportion", fill = legend_title) +
    theme_minimal()
}

ADLPlot1 <- make_adl_plot(ADLPlotData1, KOOS12DailyLivingQ1_6months, "Sit-stand difficulty",      pal_4)
ADLPlot2 <- make_adl_plot(ADLPlotData2, KOOS12DailyLivingQ2_6months, "Standing difficulty",      pal_4)
ADLPlot3 <- make_adl_plot(ADLPlotData3, KOOS12DailyLivingQ3_6months, "Vehicle egress difficulty",  pal_4)
ADLPlot4 <- make_adl_plot(ADLPlotData4, KOOS12DailyLivingQ4_6months, "Twisting difficulty",        pal_4)

# ── Combine ───────────────────────────────────────────────────────────────────
(ADLPlot1 + ADLPlot2) / (ADLPlot3 + ADLPlot4)
```


```{r}
knitr::knit_print(KneelPlot)
```



```{r}
#| label: fig-model-kneeling1
#| fig-cap: "Adjusted differences in incidence of responses for kneeling difficulty at 6months between pre and post introduction of cr insert"



# Collapse Extreme into Severe, then set reference level
GMKMaster <- GMKMaster |>
  mutate(
    Kneeling_6months = case_when(
      Kneeling_6months == "Extreme" ~ "Severe",
      .default = Kneeling_6months
    ),
    Kneeling_6months = relevel(factor(Kneeling_6months), ref = "Moderate"),
    Kneeling_6monthsAlt = case_when(
      Kneeling_6months %in% c("Moderate", "Severe") ~ "High",
      Kneeling_6months %in% c("Mild", "None") ~ "Low",
      .default = NA_character_
    ),
    Kneeling_6monthsAlt = relevel(factor(Kneeling_6monthsAlt), ref = "High")
  )

# Fit multinomial logistic regression
# Adjusting for age and sex
# The first level of categorical_outcome will be the reference category
MultiNommodel <- nnet::multinom(
  formula = Kneeling_6months ~ CRTiming + AgeAtTreatment + Sex,
  data = GMKMaster
)

# Extract tidy results with odds ratios and confidence intervals
KneelResults <- MultiNommodel |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  filter(str_detect(term, "CRTiming")) |>
  mutate(
    across(c(estimate, conf.low, conf.high), ~round(., 2)),
    p.value = round(p.value, 4)
  ) |>
  select(
    outcome = y.level,
    term,
    odds_ratio = estimate,
    ci_lower = conf.low,
    ci_upper = conf.high,
    p.value
  )

# Plot
KneelResults |>
  mutate(outcome = factor(outcome, levels = c("Severe", "None", "Moderate", "Mild"))) |>
  ggplot(aes(x = odds_ratio, y = outcome)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Outcome Level"
  ) +
  theme_bw()
```


```{r}
#| label: fig-model-kneeling2
#| fig-cap: "Adjusted differences in incidence of low kneeling difficulty at 6months relative to high difficulty between pre and post introduction of cr insert"

# Fit binomial logistic regression
# Adjusting for age and sex
# High is the reference category (odds of being Low vs High)
BinomModel <- glm(
  formula = Kneeling_6monthsAlt ~ CRTiming + AgeAtTreatment + Sex,
  data    = GMKMaster,
  family  = binomial
)

# Extract tidy results with odds ratios and confidence intervals
KneelResultsBinom <- BinomModel |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  filter(str_detect(term, "CRTiming")) |>
  mutate(
    across(c(estimate, conf.low, conf.high), ~round(., 2)),
    p.value = round(p.value, 4)
  ) |>
  select(
    term,
    odds_ratio = estimate,
    ci_lower  = conf.low,
    ci_upper  = conf.high,
    p.value
  )

# Plot
KneelResultsBinom |>
  ggplot(aes(x = odds_ratio, y = term)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Incidence Low Difficulty relative to High Difficulty"
  ) +
  theme_bw()

```

```{r}
#| label: fig-model-sittostand1
#| fig-cap: "Adjusted differences in incidence of responses for sit-to-stand difficulty at 6months between pre and post introduction of cr insert"

# ─────────────────────────────────────────────
# KOOS12 Sit-to-Stand (KOOS12DailyLivingQ1_6months)
# ─────────────────────────────────────────────


GMKMaster <- GMKMaster |>
  mutate(
    KOOS12DailyLivingQ1_6months = case_when(
      KOOS12DailyLivingQ1_6months == "Extreme" ~ "Severe",
      .default = KOOS12DailyLivingQ1_6months
    ),
    KOOS12DailyLivingQ1_6months = relevel(factor(KOOS12DailyLivingQ1_6months), ref = "Moderate"),
    KOOS12DailyLivingQ1_6monthsAlt = case_when(
      KOOS12DailyLivingQ1_6months %in% c("Moderate", "Severe") ~ "High",
      KOOS12DailyLivingQ1_6months %in% c("Mild", "None")       ~ "Low",
      .default = NA_character_
    ),
    KOOS12DailyLivingQ1_6monthsAlt = relevel(factor(KOOS12DailyLivingQ1_6monthsAlt), ref = "High")
  )

# Multinomial model
MultiNommodel_SitStand <- nnet::multinom(
  formula = KOOS12DailyLivingQ1_6months ~ CRTiming + AgeAtTreatment + Sex,
  data    = GMKMaster
)

SitStandResults <- MultiNommodel_SitStand |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  filter(str_detect(term, "CRTiming")) |>
  mutate(
    across(c(estimate, conf.low, conf.high), ~round(., 2)),
    p.value = round(p.value, 4)
  ) |>
  select(
    outcome   = y.level,
    term,
    odds_ratio = estimate,
    ci_lower   = conf.low,
    ci_upper   = conf.high,
    p.value
  )

SitStandResults |>
  mutate(outcome = factor(outcome, levels = c("Severe", "None", "Moderate", "Mild"))) |>
  ggplot(aes(x = odds_ratio, y = outcome)) +
  geom_point(size = 2) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2, orientation = "y") +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Outcome Level"
  ) +
  theme_bw()

```


```{r}
#| label: fig-model-sittostand2
#| fig-cap: "Adjusted differences in incidence of low sit-to-stand difficulty at 6months relative to high difficulty between pre and post introduction of cr insert"

BinomModel_SitStand <- glm(
  formula = KOOS12DailyLivingQ1_6monthsAlt ~ CRTiming + AgeAtTreatment + Sex,
  data    = GMKMaster,
  family  = binomial
)

SitStandResultsBinom <- BinomModel_SitStand |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  filter(str_detect(term, "CRTiming")) |>
  mutate(
    across(c(estimate, conf.low, conf.high), ~round(., 2)),
    p.value = round(p.value, 4)
  ) |>
  select(
    term,
    odds_ratio = estimate,
    ci_lower   = conf.low,
    ci_upper   = conf.high,
    p.value
  )

SitStandResultsBinom |>
  ggplot(aes(x = odds_ratio, y = term)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Incidence Low Difficulty relative to High Difficulty"
  ) +
  theme_bw()

```

```{r}
#| label: fig-model-vehicleegress1
#| fig-cap: "Adjusted differences in incidence of responses for vehicle egress difficulty at 6months between pre and post introduction of cr insert"

# ─────────────────────────────────────────────
# KOOS12 Vehicle Egress (KOOS12DailyLivingQ3_6months)
# ─────────────────────────────────────────────

GMKMaster <- GMKMaster |>
  mutate(
    KOOS12DailyLivingQ3_6months = case_when(
      KOOS12DailyLivingQ3_6months == "Extreme" ~ "Severe",
      .default = KOOS12DailyLivingQ3_6months
    ),
    KOOS12DailyLivingQ3_6months = relevel(factor(KOOS12DailyLivingQ3_6months), ref = "Moderate"),
    KOOS12DailyLivingQ3_6monthsAlt = case_when(
      KOOS12DailyLivingQ3_6months %in% c("Moderate", "Severe") ~ "High",
      KOOS12DailyLivingQ3_6months %in% c("Mild", "None")       ~ "Low",
      .default = NA_character_
    ),
    KOOS12DailyLivingQ3_6monthsAlt = relevel(factor(KOOS12DailyLivingQ3_6monthsAlt), ref = "High")
  )

# Multinomial model
MultiNommodel_VehicleEgress <- nnet::multinom(
  formula = KOOS12DailyLivingQ3_6months ~ CRTiming + AgeAtTreatment + Sex,
  data    = GMKMaster
)

VehicleEgressResults <- MultiNommodel_VehicleEgress |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  filter(str_detect(term, "CRTiming")) |>
  mutate(
    across(c(estimate, conf.low, conf.high), ~round(., 2)),
    p.value = round(p.value, 4)
  ) |>
  select(
    outcome    = y.level,
    term,
    odds_ratio = estimate,
    ci_lower   = conf.low,
    ci_upper   = conf.high,
    p.value
  )

VehicleEgressResults |>
  mutate(outcome = factor(outcome, levels = c("Severe", "None", "Moderate", "Mild"))) |>
  ggplot(aes(x = odds_ratio, y = outcome)) +
  geom_point(size = 2) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2, orientation = "y") +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Outcome Level"
  ) +
  theme_bw()

```

```{r}
#| label: fig-model-vehicleegress2
#| fig-cap: "Adjusted differences in incidence of low vehicle egress difficulty at 6months relative to high difficulty between pre and post introduction of cr insert"

BinomModel_VehicleEgress <- glm(
  formula = KOOS12DailyLivingQ3_6monthsAlt ~ CRTiming + AgeAtTreatment + Sex,
  data    = GMKMaster,
  family  = binomial
)

VehicleEgressResultsBinom <- BinomModel_VehicleEgress |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  filter(str_detect(term, "CRTiming")) |>
  mutate(
    across(c(estimate, conf.low, conf.high), ~round(., 2)),
    p.value = round(p.value, 4)
  ) |>
  select(
    term,
    odds_ratio = estimate,
    ci_lower   = conf.low,
    ci_upper   = conf.high,
    p.value
  )

VehicleEgressResultsBinom |>
  ggplot(aes(x = odds_ratio, y = term)) +
  geom_point(size = 2) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), height = 0.2, orientation = "y") +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Incidence Low Difficulty relative to High Difficulty"
  ) +
  theme_bw()
```

